\section[拗音]{\ruby{拗}{よう}\ruby{音}{おん}}\label{sec:PR仮名拗音}

Now to answer the second question: ``How do you write `jyuu?''' This is done through youon, translating to `crooked sound.' You twist and fuse two sounds together to come out with a fusion of both.

How this actually works in practice, you take any of the Ｉ-row kana and add a small Ｙ-column kana to the end of it. The effect of this is to delete that Ｉ and replace it with the appended kana. To use another animal example, cats go 「ニャンニャン」, read as ``nyan nyan.'' Below is the chart for the youon combinations, the base character chart has been ommitted, remember the sound mergers and sound shifts of the base kana while reading this.

\input{Primitives/kana/resources/DakutenHiraganaYouon}

\input{Primitives/kana/resources/DakutenKatakanaYouon}

There is however another kind of youon. Let's take a look at the name of the famous game series Final Fantasy. It's written in Japanese as 「ファイナルファンタジー」 and is read as ``fainaru fantajii''. Taking that `fu' kana and adding a small `a' kana to it allows us to delete and replace the vowel there too! This technique can apply to all vowels, and is typically used for making sounds that don't appear in the charts above. Take for example, the word ``discord.'' If you try to create a `di' kana it would be a `dzi/ji' kana, so you have to improvise and create 「ディ」 instead. Those are the most common kinds of additional youon replacements, using `fu' to generate a new Ｆ-column of the gojyuuon, and building your own actual `di' sound.